---
output: github_document
---

```{r, include = FALSE}
# This document is reproduced here from the vignette 'Devil's Hole' in the `isogeochem` package
# https://github.com/davidbajnai/isogeochem
```

```{r, include = FALSE}
#> Save the user's options and parameters
oldpar = par()
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  eval = TRUE,
  fig.path = "figures/devils_hole-",
  fig.height = 4,
  fig.width = 8,
  dpi = 150,
  out.width = "100%"
)
knitr::opts_knit$set(global.par = TRUE)
```

```{r, include = FALSE}
#> Set new parameters for this vignette.
par(mfrow = c(1, 1), mar = c(4.5, 4.5, 0.3, 0.3), bg = "white")
```

## Introduction

This vignette demonstrates the usage of `isogeochem` using a case study based on
[Bajnai et al. (2021)](https://doi.org/10.1029/2021GL093257).

#### Background:

[Winograd et al. (2006)](https://doi.org/10.1016/j.yqres.2006.06.003) acquired a *δ*<sup>18</sup>O time series from carbonates that precipitated underwater in the Devils Hole caves spanning the last ca. 570 thousand years. Now, we are interested in the variations in the *δ*<sup>18</sup>O value of the groundwater in this period. To reconstruct a groundwater *δ*<sup>18</sup>O time series from the carbonate *δ*<sup>18</sup>O values, we have to know the temperature of the groundwater. Therefore, we measured the clumped isotope values of ten calcite samples. First, we convert the measured *∆*<sub>47</sub> values to carbonate growth temperatures. Then, using the mean of the clumped isotope temperatures, we convert the carbonate *δ*<sup>18</sup>O time series to a groundwater *δ*<sup>18</sup>O time series.

## Package setup

Download, install, and load the `isogeochem` package:

```{r, include = TRUE, message = FALSE, eval = TRUE}
if (!require("isogeochem")) install.packages("isogeochem")
library("isogeochem")
```

